### 2023-04-27 figures/N3_CLVR_cycle.tex
Started results section and added preliminary figures and tables

\documentclass[landscape]{article}

\usepackage[latin1]{inputenc}
\usepackage{tikz}
\usetikzlibrary{shapes, arrows, positioning, calc, matrix}
\usepackage{xcolor}

%%%<
\usepackage{verbatim}
\usepackage[active,tightpage]{preview}
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{5pt}%
%%%>

\begin{document}
\pagestyle{empty}

% define colors
\definecolor{shaftcolor}{RGB}{255,0,0}
\definecolor{elementcolor}{RGB}{46, 187, 242}

% Define block styles

\tikzstyle{element} = [rectangle, draw, fill=elementcolor!20,
text width=9em, text centered, font=\LARGE, rounded corners, minimum height=6em]
\tikzstyle{shaft_element} = [rectangle, draw, fill=shaftcolor!20,
text width=10em, text centered, font=\LARGE, rounded corners, minimum height=4em]
\tikzstyle{heat_element} = [rectangle, draw, fill=heatcolor!20,
text width=9em, text centered, font=\LARGE, rounded corners, minimum height=6em]
\tikzstyle{connection} = [draw, -latex', ultra thick]
\tikzstyle{shaft} = [draw, line width=0.1cm, color=shaftcolor]
\tikzstyle{heat_flow} = [draw, -latex', ultra thick, color=heatcolor]

\begin{tikzpicture}[node distance = 2cm, auto]
    % Create elements
    % \node [element] (fc) {Flight Conditions};
    \node [element] (inlet) {Inlet};
    \node [element, right=4em of inlet] (fan) {Fan};
    \node [element, right=4em of fan] (splitter) {Splitter};
    \node [element, right=4em of splitter] (lpc) {Low Pressure Compressor};
    \node [element, right=4em of lpc ] (injector) {Water Injector};
    \node [element, right=4em of injector ] (hpc) {High Pressure Compressor};
    \node [element, right=4em of hpc] (burner) {Burner};
    \node [element, right=4em of burner] (hpt) {High-Pressure Turbine};
    \node [element, right=4em of hpt] (lpt) {Low-Pressure Turbine};
    \node [element, right=4em of lpt] (extractor) {Water Extractor};
    \node [element, right=4em of extractor] (corenozzle) {Core Nozzle};
    \node [element, below=4em of burner] (vrl) {Vapor Recovery Loop};
    \node [element, below=12em of burner] (bypduct) {Bypass Duct};
    \node [element, below=12em of corenozzle] (bypnozzle) {Bypass Nozzle};

    % Add shaft elements
    \node [shaft_element, above=8em of $(hpc)!0.5!(hpt)$] (hpshaft) {High Pressure Shaft};
    \node [shaft_element, above=16em of $(lpc)!0.5!(lpt)$] (lpshaft) {Low Pressure Shaft};
    \node [shaft_element, above=16em of $(fan)!0.28!(lpc)$] (fanshaft) {Fan Shaft};
    \node [shaft_element, above=16em of $(fan)!0.72!(lpc)$] (gearbox) {Gearbox};

    % Draw element connections
    % \path [connection] (fc) -- (inlet);
    \path [connection] (inlet) -- (fan);
    \path [connection] (fan) -- (splitter);
    \path [connection] (splitter) -- (lpc);
    \path [connection] (splitter) |- (bypduct);
    \path [connection] (bypduct) -- (bypnozzle);
    \path [connection] (lpc) -- (injector);
    \path [connection] (injector) -- (hpc);
    \path [connection] (hpc) -- (burner);
    \path [connection] (burner) -- (hpt);
    \path [connection] (hpt) -- (lpt);
    \path [connection] (lpt) -- (extractor);
    \path [connection] (extractor) -- (corenozzle);
    \path [connection] (extractor) |- (vrl);
    \path [connection] (vrl) -| (injector);

    % Draw Shaft Connections
    \path [shaft] (fan) |- (fanshaft) -- (gearbox);
    \path [shaft] (gearbox) -| (lpc);
    \path [shaft] (lpc) |- (lpshaft) -| (lpt);
    \path [shaft] (hpc) |- (hpshaft) -| (hpt);

    % \matrix[draw, ultra thick, rounded corners, matrix of nodes, text width=30em, font=\LARGE, row sep=3mm, below=25em of current page.south west, anchor=west] (key) {
    %     $\bullet$ Black arrows are flow connections    \\
    %     $\bullet$ Red lines are mechanical connections \\
    %     $\bullet$ Blue boxes are cycle elements        \\
    %     $\bullet$ Red boxes are shaft elements         \\
    % };

\end{tikzpicture}

\end{document}