### 2023-05-01 figures/N3_cycle.tex
Fixed some formating and scentence issues

\documentclass[landscape]{article}

\usepackage[latin1]{inputenc}
\usepackage{tikz}
\usetikzlibrary{shapes, arrows, positioning, calc, matrix}
\usepackage{xcolor}

%%%<
\usepackage{verbatim}
\usepackage[active,tightpage]{preview}
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{5pt}%
%%%>

\begin{document}
\pagestyle{empty}

% define colors
\definecolor{shaftcolor}{RGB}{255,0,0}
\definecolor{elementcolor}{RGB}{46, 187, 242}

% Define block styles

\tikzstyle{element} = [rectangle, draw, fill=elementcolor!20,
text width=9em, text centered, font=\LARGE, rounded corners, minimum height=6em]
\tikzstyle{shaft_element} = [rectangle, draw, fill=shaftcolor!20,
text width=10em, text centered, font=\LARGE, rounded corners, minimum height=4em]
\tikzstyle{heat_element} = [rectangle, draw, fill=heatcolor!20,
text width=9em, text centered, font=\LARGE, rounded corners, minimum height=6em]
\tikzstyle{connection} = [draw, -latex', ultra thick]
\tikzstyle{shaft} = [draw, line width=0.1cm, color=shaftcolor]
\tikzstyle{station_label} = [circle, draw, text width=2em, text centered, font=\LARGE]
\tikzstyle{line} = [draw]

\begin{tikzpicture}[node distance = 2cm, auto]
    % Create elements
    % \node [element] (fc) {Flight Conditions};
    \node [element] (inlet) {Inlet};
    \node [element, right=4em of inlet] (fan) {Fan};
    \node [element, right=4em of fan] (splitter) {Splitter};
    \node [element, right=4em of splitter] (lpc) {Low Pressure Compressor};
    \node [element, right=4em of lpc ] (hpc) {High Pressure Compressor};
    \node [element, right=4em of hpc] (burner) {Burner};
    \node [element, right=4em of burner] (hpt) {High-Pressure Turbine};
    \node [element, right=4em of hpt] (lpt) {Low-Pressure Turbine};
    \node [element, right=4em of lpt] (corenozzle) {Core Nozzle};
    \node [element, below=8em of burner] (bypduct) {Bypass Duct};
    \node [element, below=8em of corenozzle] (bypnozzle) {Bypass Nozzle};

    % Add shaft elements
    \node [shaft_element, above=8em of $(hpc)!0.5!(hpt)$] (hpshaft) {High Pressure Shaft};
    \node [shaft_element, above=16em of $(lpc)!0.5!(lpt)$] (lpshaft) {Low Pressure Shaft};
    \node [shaft_element, above=16em of $(fan)!0.28!(lpc)$] (fanshaft) {Fan Shaft};
    \node [shaft_element, above=16em of $(fan)!0.72!(lpc)$] (gearbox) {Gearbox};

    % Create station labels
    % \node [station_label, below=8em of inlet.west] (1) {1};
    \node [station_label, below=6em of fan.west] (1) {1};
    % \node [station_label, below=8em of splitter.west] (13) {13};
    \node [station_label, below=6em of lpc.west] (2) {2};
    \node [station_label, below=6em of hpc.west] (25) {25};
    \node [station_label, below=6em of burner.west] (3) {3};
    \node [station_label, below=6em of hpt.west] (4) {4};
    \node [station_label, below=6em of lpt.west] (45) {45};
    \node [station_label, below=6em of corenozzle.west] (5) {5};
    \node [station_label, below=6em of bypnozzle.west] (17) {17};

    % Draw element connections
    % \path [connection] (fc) -- (inlet);
    \path [connection] (inlet) -- (fan);
    \path [connection] (fan) -- (splitter);
    \path [connection] (splitter) -- (lpc);
    \path [connection] (splitter) |- (bypduct);
    \path [connection] (bypduct) -- (bypnozzle);
    \path [connection] (lpc) -- (hpc);
    \path [connection] (hpc) -- (burner);
    \path [connection] (burner) -- (hpt);
    \path [connection] (hpt) -- (lpt);
    \path [connection] (lpt) -- (corenozzle);

    % Draw station label connections
    % \path [line] (inlet.west) -- (1.north);
    \path [line] (fan.west) -- (1.north);
    % \path [line] (splitter.west) -- (13.north);
    \path [line] (lpc.west) -- (2.north);
    \path [line] (hpc.west) -- (25.north);
    \path [line] (burner.west) -- (3.north);
    \path [line] (hpt.west) -- (4.north);
    \path [line] (lpt.west) -- (45.north);
    \path [line] (corenozzle.west) -- (5.north);
    \path [line] (bypnozzle.west) -- (17.north);

    % Draw Shaft Connections
    \path [shaft] (fan) |- (fanshaft) -- (gearbox);
    \path [shaft] (gearbox) -| (lpc);
    \path [shaft] (lpc) |- (lpshaft) -| (lpt);
    \path [shaft] (hpc) |- (hpshaft) -| (hpt);

    % \matrix[draw, ultra thick, rounded corners, matrix of nodes, text width=30em, font=\LARGE, row sep=3mm, below=25em of current page.south west, anchor=west] (key) {
    %     $\bullet$ Black arrows are flow connections    \\
    %     $\bullet$ Red lines are mechanical connections \\
    %     $\bullet$ Blue boxes are cycle elements        \\
    %     $\bullet$ Red boxes are shaft elements         \\
    % };

\end{tikzpicture}

\end{document}